---
title: "Mini-Project 2"
author: "Angelica Estrada"
date: "October 31, 2017"
output: html_document
---

<!-- GitHub Steps: Pull, Stage, Commit, Push -->

## Loading the data

This project is so interesting and relevant to my life!

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
library(tidyverse)
library(ggplot2)
library(plotly)
packageVersion('plotly')
library(scales)
```

```{r} 
library(tidyverse)
glimpse(house_elections)
```

```{r}
glimpse(candidates)
```

```{r}
glimpse(committees)
```

```{r}
glimpse(contributions)
```

<!-- SKIP THIS
```{r}
# Create candidates data set for only year 2011
can2011 <- candidates %>%
  filter(cand_election_yr == 2011)

# Repeat for year 2012
can2012 <- candidates %>%
  filter(cand_election_yr == 2012)

```

```{r}
# Check for duplicated candidates in `can2011`
any(duplicated(can2011$cand_name))

# Check for duplicated candidates in `can2012`
any(duplicated(can2012$cand_name))

```
-->

## FOR YEAR 2011
```{r}
# Find number of candidates who ran for P,H,S
can_num2011 <- candidates %>%
  filter(cand_election_yr == 2011) %>%
  group_by(cand_office) %>%
  summarise(num = n())
can_num2011
```

```{r}
# Create dataset of candidates who ran for President
pres2011 <- candidates %>%
  filter(cand_election_yr == 2011 & cand_office == "P") 

# Create dataset of candidates who ran for House
house2011 <- candidates %>%
  filter(cand_election_yr == 2011 & cand_office == "H") 
 
# Check duplicates
any(duplicated(pres2011$cand_name))
any(duplicated(house2011$cand_name))

```

## FOR YEAR 2012

```{r}
# Find number of candidates who ran for P,H,S
can_num2012 <- candidates %>%
  filter(cand_election_yr == 2012) %>%
  group_by(cand_office) %>%
  summarise(num = n())
can_num2012
```


```{r}

# Create dataset of candidates who ran for President
pres2012 <- candidates %>%
  filter(cand_election_yr == 2012 & cand_office == "P") 
 
# Create dataset of candidates who ran for House
house2012 <- candidates %>%
  filter(cand_election_yr == 2012 & cand_office == "H") 
 
# Create dataset of candidates who ran for Senate
sen2012 <- candidates %>%
  filter(cand_election_yr == 2012 & cand_office == "S") 
 
# Check duplicates
any(duplicated(pres2012$cand_name))
any(duplicated(house2012$cand_name))
any(duplicated(sen2012$cand_name))
```
 
```{r}
# Check the duplicate in house2012
d <- house2012$cand_name[duplicated(house2012$cand_name)] 
d
```

```{r}
hancock <- house2012 %>%
  filter(cand_name == "HANCOCK, DAVID")

```

```{r}
dupes <- house2012 %>%
  filter(cand_name == "HANCOCK, DAVID" |
           cand_name == "AVERSA, ALAN" |
           cand_name == "BIBBS, JAMES" |
           cand_name == "LEWIS, DAVID" |
           cand_name == "FOSTER, JAMES" |
           cand_name == "SANCHEZ, DAVID" | 
           cand_name == "BRODHEAD, STEPHAN ANDREW" |
           cand_name == "HATHCOX, VALINDA")

dupes <- dupes %>%
  arrange(cand_name)

```

<!-- IGNORE (TRIED MAKING A FUNCTION INSTEAD OF MANUALLY CODING ABOVE)
```{r}
# Create duplicate function
dup_fun2012 <- function(name){
  house2012 %>%
  filter(cand_name == name)
}
```

```{r}
# Use lapply
lapply(d, dup_fun2012)
```
--> 
 
 Clean up of data, only include 2012? 
 
```{r}
cleancontributions <- contributions %>%
  select(cmte_id, name, city, state, zip_code, transaction_amt, cand_id, tran_id, transaction_type) %>%
  filter( transaction_type == "24K")
cleancandidates <- candidates %>%
  filter( cand_election_yr == 2012, cand_status == "C") %>%
  select(cand_id, cand_name, cand_party_affiliation, cand_election_yr, cand_office_state, cand_office, cand_status, cand_city, cand_state, cand_zip)
```

Join the two datasets by cand_id:

```{r}
candcontjoin <- cleancontributions %>%
  full_join( cleancandidates, by = "cand_id") %>%
  filter( cand_party_affiliation == "DEM" | cand_party_affiliation == "REP")
```


Then add the committee name based off of Committee ID:
```{r}
committeename <- committees %>%
  select(cmte_id, cmte_name)
```

```{r}
committeecontributionlist <- candcontjoin %>%
  left_join( committeename, by = "cmte_id")
```


Group by number of donations to each committee:

```{r}
committeegroup <- committeecontributionlist %>%
  group_by( cmte_name ) %>%
  summarize( countofcontributions = n(), totaltransaction = sum(transaction_amt))
committeegroup[is.na(committeegroup)] <- 0
```

```{r}
demcommittee <- committeecontributionlist %>%
  filter(cand_party_affiliation == "DEM") %>%
  group_by(cmte_name)%>%
  summarize( countofdem = n())
```

```{r}
repcommittee <- committeecontributionlist %>%
  filter(cand_party_affiliation == "REP") %>%
  group_by(cmte_name)%>%
  summarize( countofrep = n())
```


```{r}
committeebyparty <- demcommittee %>%
  full_join( repcommittee, by = "cmte_name") 
committeebyparty[is.na(committeebyparty)] <- 0
```

Check to make sure the joining is working properly:

```{r}
committeebypartycheck <-committeebyparty %>%
  full_join( committeegroup, by = "cmte_name") %>%
  mutate( check = countofcontributions - (countofdem + countofrep)) 
```


Find the percent of total donors that were either Rep or Dem:

```{r}
committeebypartyfinal <-committeebyparty %>%
  full_join( committeegroup, by = "cmte_name") %>%
  mutate( Democratic = (countofdem / countofcontributions) * 100) %>%
  mutate( Republican = (countofrep / countofcontributions) * 100) %>%
  mutate( party = ifelse( Democratic > Republican, "D", "R")) %>%
  mutate( party = ifelse( Democratic == Republican, "Neutral", party)) %>%
  mutate( totalperc = Democratic + Republican) %>%
  rename( Committee = cmte_name)
```

Largest Committees:
```{r}
largestcommittee <- committeebypartyfinal %>%
  arrange(desc(totaltransaction))%>%
  head(10)
```

       
```{r}
largestcommitteeclean <- largestcommittee %>%
  select(Committee, Republican, Democratic)
mlargestcommittee <- melt(largestcommitteeclean, id= c("Committee"))
p <- ggplot(mlargestcommittee, aes( x= Committee, value, fill = variable)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values=c("firebrick","royalblue4"),
                    name="Party\nAffliation") +
  scale_y_continuous(labels = percent) +
  theme(axis.text.x=element_blank()) +
  labs(title= "Party Affiliation of Largest US Committees", y ="Percentage", x = "Top 10 Largest Committees")
p <- ggplotly(p)
p
```





